An Empirical Analysis of the Internet Engineering Task Force with Computational Methods

### 2024-10-03 paper.tex
Briefly summarise some results from our Frontiers paper

\pb{Changes in Use of Language with Changes in Role:}
Finally, we look at how linguistic behaviour of participants changes as
they gain influence in the IETF.
We consider at participants who went from low to high influence over time:
individuals who had a \emph{centrality-based} influence below the 50th
percentile when they joined the IETF, and reached the top 10th percentile
at some point.

@@

%..................................................................................................
\pb{Discussion:}
To better understand these LIWC categories and what kind of words play a
role in the behaviour of individual categories, we calculate the frequency
of words in each LIWC category as they appear in the emails. 
Next, we consider the top 30 most frequent words in each LIWC category and
perform regression analysis on centrality-based influence for participants,
but using only these 30 words as features to generate the participant
representation. We conducted this experiment separately for each LIWC
category that was significant in the first experiment.

@@

%..................................................................................................
\pb{Summary:}
Using two aspects of influence, based on centrality in the email social
graph and based on organisation role, we were able to unfold several traits
that differentiate influential participants from others. Many of our
findings seem corroborated by studies in organisational theory. 
We observed that influential people exhibit more collaborative and
community-oriented traits, and also stronger signs of engagement in
discussions. We also observed that as people go on to become influential
participants, they evolve in their communication and are seen to be more
engaging and descriptive in their linguistic style.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

% Frontiers in Psychology paper
% - Use of Language, discussing how it shifts as people become more senior

% LEDA
% - a taxonomy of dialogue acts (DAs) and a labeled dataset of emails

>>>>

\pb{Changes in Language with Changes in Role:}
Finally, we look at how linguistic behaviour of participants changes as
they gain influence in the IETF.
We consider at participants who went from low to high influence over time:
individuals who had a \emph{centrality-based} influence below the 50th
percentile when they joined the IETF, and reached the top 10th percentile
at some point.

@@

%..................................................................................................
% The following is from our ACL 2023 Tracing Linguistic Markers of
% Influence paper, Section 4.2.
\pb{What Kinds of Language are Used?}
To better understand these LIWC categories and what kind of words play a
role in the behaviour of individual categories, we calculate the frequency
of words in each LIWC category as they appear in the emails. 
Next, we consider the top 30 most frequent words in each LIWC category and
perform regression analysis on centrality-based influence for participants,
but using only these 30 words as features to generate the participant
representation. We conducted this experiment separately for each LIWC
category that was significant in the first experiment.

@@

%..................................................................................................
% The following is from our Frontiers in Psychology paper:

\pb{Use of Sensitive Language:}
We briefly consider the use of sensitive language, relating to politics,
race, and religion, in IETF email communications.

Using email data for 2019, the last year prior to the pandemic, we
categorise the emails according to the role of their sender, based
on the relative social power of the sender within a working group:
\emph{working group chair} for emails sent by the current chair of the
working group to which the email was sent;
\emph{other chair} for emails sent by the chair of a different working
group; and
\emph{regular participant} for emails sent by people who were not, and had
not been at the time of sampling, a workgroup chair.
Due to their small numbers, we exclude area directors.

We use three LIWC categories to infer the presence of potentially sensitive
language in the emails: 
\liwc{Politics}---words commonly used in political discussions;
\liwc{Ethnicity}---words that identify national, regional, linguistic,
  ethnic, or racial identities; and
\liwc{Religion}--use of religious words.
Each email is coded according to the working group and role of the sender.
The emails are pre-processed to remove subject lines, headers, and any
embedded quotes from previous emails, yielding 101,857 texts by 2,300
individuals across 176 group mailing lists.  Each text is assigned a score
on each of the three language categories using the LIWC software and the
resulting scores are averaged for each individual sender for statistical
analysis.

Results are shown in Figure \ref{fig:sensitive}. The consistent pattern
is that people with organizational responsibilities (working group chair,
other chair) are less likely to use words potentially connected with
sensitive topics than regular participants without organizational
responsibilities, even though they are all part of the same discussions
on the same email lists.
Averaging across the two working group chair categories, the means indicate
that people in positions of social power are 3.2 times less likely to use
sensitive language overall than those who are not (by category:
\liwc{Politics} 2.0 times; \liwc{Religion} 2.3 times; \liwc{Ethnicity} 5.5
times).
It is worth noting, however, that in absolute terms potentially sensitive
language use is rare in this dataset, as might be expected for primarily
work-related communication. The modal LIWC score in both raw data and the
averages scores is zero

\begin{figure*}[t]
  \centering
  \includegraphics[width=\figureWidthTwoColumn]{figures-prev/frontiers/sensitive.pdf}
  \caption{
    Patterns of language use by organizational role. Estimated marginal
    means for the LIWC categories: (A) Politics, (B) Ethnicity, and (C)
    Religion. Error bars show 95% confidence intervals.
  }
  \label{fig:sensitive}
\end{figure*}

%..................................................................................................
% The following is from our ACL 2023 Tracing Linguistic Markers of
% Influence paper
\pb{Summary:}
Using two aspects of influence, based on centrality in the email social
graph and based on organisation role, we were able to unfold several
traits that differentiate influential participants from others. Many
of our findings seem corroborated by studies in organisational theory.
We observed that influential people exhibit more collaborative and
community-oriented traits, show stronger signs of engagement in
discussions, and are systematically less likely to use potentially
sensitive language.
We also observed that as people go on to become influential participants,
they evolve in their communication and are seen to be more engaging and
descriptive in their linguistic style.
